Kids SUMs Summary

### 2020-02-14 exposure_initial_analysis.Rmd
exposure analysis - simple associations

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

@@

# Combine datasets
```{r}
pm_home_mean <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  #filter(pm < 500) %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  group_by(area, home) %>% 
  summarize(mean_pm = mean(pm, na.rm = TRUE))

pm_demographics <- pm_home_mean %>% 
  left_join(demographics_clean, by = c("area", "home")) %>% 
  filter(winter_id == 1) %>% 
  mutate(income = factor(income,
                         levels = c("Less than $20,000", "$20,000 to $29,999",
                                    "$30,000 to $39,999", "$40,000 to $49,999",
                                    "$50,000 to $74,999", "$75,000 to $99,999",
                                    "$100,000 or more"))) %>% 
  mutate(education = factor(education,
                            levels = c("Less than high school", 
                                       "High school diploma or GED",
                                       "Some college",
                                       "College degree")))

pm_home_char <- pm_home_mean %>% 
  left_join(home_char_clean, by = c("area", "home")) %>% 
  filter(winter_id == 1) %>% 
  mutate(stove_age = factor(stove_age,
                         levels = c("0-5 years old", "6-10 years old",
                                    "11-15 years old", "16 + years old",
                                    "Do not know")))

pm_home_char <- pm_home_mean %>% 
  left_join(home_char_clean, by = c("area", "home")) %>% 
  filter(winter_id == 1) %>% 
  mutate(stove_age = factor(stove_age,
                         levels = c("0-5 years old", "6-10 years old",
                                    "11-15 years old", "16 + years old",
                                    "Do not know")))
```

# Simple associations between PM and covariates
```{r}
pm_demographics_results <- lm(mean_pm ~ education, data = pm_demographics)
summary(pm_demographics_results)

pm_home_char_results <- lm(mean_pm ~ stove_age, data = pm_home_char)
summary(pm_home_char_results)
```

>>>>

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(lme4)
library(lmerTest)
```

@@

# Combine datasets
```{r}
# Mean PM for each placebo home over winter 1
pm_home_mean <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  #filter(pm < 500) %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  group_by(area, home) %>% 
  summarize(mean_pm = mean(pm, na.rm = TRUE))

# Join demographics with mean PM and format for analysis
pm_demographics <- pm_home_mean %>% 
  left_join(demographics_clean, by = c("area", "home")) %>% 
  filter(winter_id == 1) %>% 
  mutate(income = factor(income,
                         levels = c("Less than $20,000", "$20,000 to $29,999",
                                    "$30,000 to $39,999", "$40,000 to $49,999",
                                    "$50,000 to $74,999", "$75,000 to $99,999",
                                    "$100,000 or more"))) %>% 
  mutate(education = factor(education,
                            levels = c("Less than high school", 
                                       "High school diploma or GED",
                                       "Some college",
                                       "College degree")))

# Join home characteristics with mean PM and format for analysis
pm_home_char <- pm_home_mean %>% 
  left_join(home_char_clean, by = c("area", "home")) %>% 
  filter(winter_id == 1) %>% 
  mutate(stove_age = factor(stove_age,
                         levels = c("0-5 years old", "6-10 years old",
                                    "11-15 years old", "16 + years old",
                                    "Do not know")),
         stove_cert = factor(stove_cert,
                             levels = c("Yes", "No", "Do not know")),
         home_sqft_cat = cut(home_sqft, breaks = c(0, 500, 1000, 2000, 3000, 5000),
                         labels = c("<500", "500-999", "1000-1999", 
                                    "2000-2999", "3000+"),
                         right = FALSE),
         home_floors = factor(home_floors, levels = c(1,2,3)),
         home_bedrooms_cat = cut(home_bedrooms, breaks = c(0, 2, 4, 10),
                                 labels = c("0 or 1", "2 or 3", "4+"),
                                 right = FALSE),
         home_windows_cat = cut(home_windows, breaks = c(0, 5, 10, 20, 30, 50),
                                labels = c("0-5", "6-10", "11-20", "21-30", "30+")))

# Join stove use with mean PM and format for analysis
pm_stove_use <- pm_home_mean %>% 
  left_join(stove_use_clean, by = c("area", "home")) %>% 
  filter(winter_id == 1) %>% 
  mutate(burn_level = factor(burn_level,
                         levels = c("No burning", "Light burning",
                                    "Average burning", "Heavy burning"),
                         labels = c("No burning", "Light", "Average", "Heavy"))) %>% 
  mutate(wood_cords = if_else(wood_cords > 10, 10, wood_cords))

# Join demographics with mean PM and format for analysis
stove_grades <- stove_grades_clean %>% 
  arrange(area, home, winter_id) %>% 
  group_by(home) %>% 
  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade))

pm_stove_grades <- pm_home_mean %>% 
  left_join(stove_grades, by = c("area", "home")) %>% 
  filter(winter_id == 1)

# Join home activity with mean PM and format for analysis
pm_home_act <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  #filter(pm < 500) %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  arrange(area, home, sampling_day) %>% 
  group_by(area, home, sampling_day) %>% 
  summarize(mean_pm = mean(pm, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(home_act_day = as.factor(sampling_day)) %>% 
  arrange(area, home, home_act_day) %>% 
  left_join(home_act_clean, by = c("area", "home", "home_act_day")) %>% 
  filter(winter_id == 1)

# Join moisture content with mean PM and format for analyzis
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home, first_datetime_new) %>% 
  separate(first_datetime_new, c("first_date", "first_time"), sep = " ", remove = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home)

moisture_filtered <- moisture_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>%
  arrange(area, home)

moisture_filtered_time <- moisture_filtered %>% 
  # fix wrong date
  mutate(moisture_date = as.character(moisture_date),
         moisture_date = if_else(home == "CH220" & moisture_date == "2014-04-10",
                                 "2017-04-10", moisture_date),
         moisture_date = ymd(moisture_date)) %>% 
  mutate(moisture_date = ymd(moisture_date),
         first_date = ymd(first_date),
         time_diff = moisture_date - first_date,
         time_diff = as.numeric(abs(time_diff))) %>% 
  group_by(home) %>% 
  mutate(time_diff_rank = if_else(time_diff == min(time_diff, na.rm = TRUE), 1, 0)) %>%
  ungroup() %>% 
  filter(time_diff_rank == 1)

pm_moisture <- pm_home_mean %>% 
  left_join(moisture_filtered_time, by = c("area", "home")) %>% 
  filter(winter_id == 1) %>% 
  mutate(moisture_ave_5 = moisture_ave/5) %>% 
  filter(!is.na(moisture_ave_5))
```

# Simple associations between PM and covariates
```{r}
# *income*, *education*
pm_demographics_results <- lm(mean_pm ~ education, data = pm_demographics)
summary(pm_demographics_results)

# stove_age, stove_cert, *home_sqft_cat*, *secondary_heat*, *home_floors*
# home_pets, home_furry, *home_windows_cat*, *home_bedrooms_cat*, *residents_smoke*
pm_home_char_results <- lm(mean_pm ~ home_bedrooms_cat, data = pm_home_char)
summary(pm_home_char_results)
plot(pm_home_char_results)

# burn_level, wood_collect, wood_cords
pm_stove_use_results <- lm(mean_pm ~ wood_collect, data = pm_stove_use)
summary(pm_stove_use_results)

pm_stove_use_plot <- pm_stove_use %>% 
  ggplot() + 
    geom_boxplot(aes(burn_level, mean_pm))
pm_stove_use_plot

# *stove_grade*
pm_stove_grades_results <- lm(mean_pm ~ stove_grade, data = pm_stove_grades)
summary(pm_stove_grades_results)
anova(pm_stove_grades_results)

pm_stove_grades_plot <- pm_stove_grades %>% 
  ggplot() + 
    geom_boxplot(aes(stove_grade, mean_pm))
pm_stove_grades_plot

# home_act_sweep, home_act_windows, home_act_door, *home_act_smoking*
# home_act_incense, home_act_candle
pm_home_act_results <- lmer(mean_pm ~ home_act_sweep + (1 | home), data = pm_home_act)
summary(pm_home_act_results)

# *moisture_ave_5*
pm_moisture_results <- lm(mean_pm ~ moisture_ave_5, data = pm_moisture)
summary(pm_moisture_results)
```

### 2020-02-14 sums_summary.Rmd
exposure analysis - simple associations

# Define events by rate of change in temperature
```{r}
sums_temp_change <- sums_clean %>% 
  filter(winter_id == 1) %>% 
  group_by(area, home) %>% 
  mutate(time_diff = lead(datetime_sums) - datetime_sums) %>% 
  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
         lead_temp = if_else(time_diff > 25, temp_c_sums, lead_temp),
         temp_diff = lead_temp - temp_c_sums,
         temp_diff_av = abs(lead_temp - temp_c_sums),
         temp_diff_check = if_else(temp_diff >= 5, 1, 0),
         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
                              temp_diff_check)) %>% 
  ungroup() %>% 
  arrange(area, home, datetime_sums)

summary_by_area <- sums_temp_change %>% 
  group_by(area, home, date_sums) %>% 
  summarize("events" = sum(heat_event, na.rm = TRUE)) %>% 
  group_by(area) %>% 
  summarize("Mean events" = mean(events),
            "SD events" = sd(events),
            "Min events" = min(events),
            "Median events" = median(events),
            "Max events" = max(events))
summary_by_area

temp_diff_by_area <- sums_temp_change %>% 
  mutate(time_diff = as.numeric(time_diff)) %>% 
  filter(time_diff >= 5 & time_diff < 30) %>% 
  group_by(area) %>% 
  summarize("mean temp diff" = mean(temp_diff_av, na.rm = TRUE),
            "mean time diff" = mean(time_diff, na.rm = TRUE))
temp_diff_by_area
```

>>>>

# Define events by rate of change in temperature
```{r}
sums_temp_change <- sums_clean %>% 
  filter(winter_id == 1) %>% 
  #filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(time_diff = lead(datetime_sums) - datetime_sums) %>% 
  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
         lead_temp = if_else(time_diff > 25, temp_c_sums, lead_temp),
         temp_diff = lead_temp - temp_c_sums,
         temp_diff_av = abs(lead_temp - temp_c_sums),
         # change temp in the following line
         temp_diff_check = if_else(temp_diff >= 2, 1, 0),
         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
                              temp_diff_check)) %>% 
  ungroup() %>% 
  arrange(area, home, datetime_sums)

summary_by_area <- sums_temp_change %>% 
  group_by(area, treatment, home, date_sums) %>% 
  summarize("events" = sum(heat_event, na.rm = TRUE)) %>% 
  group_by(area, treatment) %>% 
  summarize("Mean events" = mean(events),
            "SD events" = sd(events),
            "Min events" = min(events),
            "25%" = quantile(events, 0.25),
            "Median events" = median(events),
            "75%" = quantile(events, 0.75),
            "Max events" = max(events))
summary_by_area

temp_diff_by_area <- sums_temp_change %>% 
  mutate(time_diff = as.numeric(time_diff)) %>% 
  filter(time_diff >= 5 & time_diff < 30) %>% 
  group_by(area, treatment) %>% 
  summarize("mean temp diff" = mean(temp_diff_av, na.rm = TRUE),
            "mean time diff" = mean(time_diff, na.rm = TRUE))
temp_diff_by_area
```
